---
title: "Project1"
author: "Yahui Li"
date: "2020/1/13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Step1 Read Data

```{r read data, echo = FALSE, message = FALSE}
#install.packages("AER")
library(AER)
data("STAR")
```

We will only examine the math scores in 1st grade in this project.
```{r}
data <- data.frame(star1 = STAR$star1, math1 = STAR$math1)

sapply(data,class)
sapply(data,summary)
```

```{r}
data.star1.na <- data[is.na(data$star1),]
all(is.na(data.star1.na$math1))
```

Which shows that the math score has not been recorded if class type is not recorded.
So we can remove the data where star1 is NA.

One of the way to deal with NA in math1 is to remove them
```{r}
data_remove_na <- na.omit(data[-is.na(data$star1),])
```
```{r}
pie(table(data_remove_na$star1),main = "pie chart of STAR class type")
tapply(data_remove_na$math1, data_remove_na$star1,summary)
boxplot(data$math1~data$star1,main = "box plot of math score in different class", 
        xlab = "STAR class type", ylab = "math score", col = c("white", "skyblue", "pink"))
```

From the result
For mean, small > regular+aide > regular
From all quantile information, small > regular+aide > regular
For min, small > other two; For max, they are the same

something interesting: there are only some certain scores like 601 612 627 653 676.

